\LARGE \bf Energy-Aware Dynamic Mission Planning Algorithm for UAVs

### 2020-10-14 icra-2021.tex
fixing text (2)

The overall plan is the union of all the missions. If for simplicity the system is sampled discrete-time
\begin{equation}
  \mathcal{M}:=\bigcup_{i\in[t_f]}{\mathcal{M}_i},
\end{equation}
the algorithm inputs $\mathcal{M}$ and outputs the position and the instantaneous energy consumption, while it adapts the controls--an action performed evolving the mission state.

@@

The control $\mathbf{u}$ along with the input matrix
\begin{equation}\label{eq:state-control}\begin{split}  
  \mathbf{u}_k&=\begin{bmatrix}g(\mathbf{s}_k)-g(\mathbf{s}_{k-1}) & \mathbf{c}_k-\mathbf{c}_{k-1}\end{bmatrix}^T,\\
  B&=\left[\begin{array}{cccc}
    1& \omega_{k,1}& \cdots& \omega_{k,\rho}  \\
     &            0&       &  \\
     &             & \ddots&  \\
     &             &       & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{s}_k$ is defined in Subsection~\ref{sec:computations-model}, $\mathbf{u}_k\in\mathbb{R}^l$ is the control given $l:=1+\rho$, $\mathbf{u}_{0}=\begin{bmatrix}0 & \cdots & 0\end{bmatrix}^T$ and $B\in\mathbb{R}^{j\times l}$. Moreover, the first item is one, while the others on the first row are gain factors $\omega_{k}\in\mathbb{R}$, quantifying the contribution of a given adjustment to the instantaneous energy.

@@

In the case of the system behaving ideally (i.e., with no uncertainty), we expect a set of states evolving accordingly to their outputs. Such observation is summarized in the following Lemma.

\begin{lem}\label{lem:state-vs-energy}
  Suppose the system of Equation~(\ref{eq:state-perf}) evolves with no uncertainty ($\mathbf{w}=\mathbf{0},v=0$) and Assumption~\ref{assm:periodic} holds. Given two time instants $k_1,k_2\in[t_f]$
  \begin{equation}
    \|\mathbf{q}_{k_1}\|\geq\|\mathbf{q}_{k_2}\|\iff y_{k_1}\geq y_{k_2}.
  \end{equation}
\end{lem}
\begin{proof}
  \begin{center}\vspace{.2cm}``\emph{The easy proof is trivial and is left as an exercise to the reader :P}''\vspace{.6cm}\end{center}
\end{proof}

@@

The mission $\mathcal{M}$ contains a set of ordered lists with $\sigma$ tasks each (recall Definition~\ref{def:mission}). These tasks are in fact simulated by $\sigma$ ROS nodes $\Psi(\mathbf{s}_k):=\left(\psi_1(s_{k,1}),\dots,\psi_\sigma(s_{k,\sigma})\right)$. Let us define the computations
\begin{equation}\label{eq:qos-def}
  \mathbf{s}_k:=\left\{s_{k,1},\dots ,s_{k,\sigma}\mid s_i\in\mathbb{S}_{k,i}\,\forall i\in[\sigma]\right\},
\end{equation}
where $s_{k,i}:\mathbb{Z}_{\geq 0}\rightarrow\mathbb{Z}_{\geq 0}$ returns the $i$-th computation, and $\mathbf{s}_k\in\mathbb{Z}_{\geq 0}^\sigma$ the set of $\sigma$ computations at time $k$.

@@

\begin{center}\vspace{.2cm}---\emph{from here on work in progress}---\vspace{.6cm}\end{center}

\section{Algorithm}
\label{sec:algo}

@@

Given mission starting point $\mathcal{M}_0$, the algorithm selects the $\mathbf{s}_k$ and $\mathbf{c}_k$ (computations and adjustments) and produce a valid state evolution
\begin{equation}\begin{split}
  \mathbf{u}:=\{(\mathbf{p}_k,\mathbf{c}_k,\mathbf{s}_k)\mid&(\varphi(\mathbf{p}_k,\mathbf{c}_k),\Psi(\mathbf{s_k}))\in\mathcal{M}_k\Longrightarrow\\
  &(\varphi(\mathbf{p}_{k+1},\mathbf{c}_{k+1}),\Psi(\mathbf{s}_{k+1}))\in\mathcal{M}_{k+1}\\
  {}&\forall k\in[t_f-1]\}.
\end{split}\end{equation}

Let us proof formally an important finding from Section~\ref{sec:model} extensively used in the algorithm.

% [Ad:] I am actually not entirely sure if the thm should be here (and if it makes sense). Aka proving formally that what we do is actually what we expect; so perhaps it does make sense but I would like your opinion before spending time on the demonstrations... Same for the lemma.
% [Ad:] And here I had the doubt if to place the thm just straight after the lemma or here. I places it here as this is actually the main contribution -> the energy model working. Does it make any sense?

\begin{thm}\label{thm:state-vs-energy}
  Consider a continuously differentiable function $\varphi_k:\mathbb{R}^2\times\mathbb{R}^{\sigma}\rightarrow\mathbb{R}$ at a time instant $k\in\mathbb{T}$.
  Assume Assumption~\ref{assm:periodic} holds, the robots is free to move in $\mathcal{P}$ defined in~(\ref{eq:area}), and is following $\varphi$ with the direction $\dot{\mathbf{p}}_d$ defined in~\ref{eq:pd}. Likewise in Lemma~\ref{lem:state-vs-energy}, the model behaves ideally. 
  Then, the instantaneous energy consumption is a linear combination of the state
  \begin{equation}
    y_k=C\mathbf{q}_k=\sum_{n=0}^r{\alpha_n},
  \end{equation}
  where $\alpha_n\in\mathbf{q}_k$ are the $r+1$ state's components at $k$ with $r$ being a preassigned arbitrary order from~(\ref{eq:fourier}), and $C$ is described in Equation~(\ref{eq:state-details}).
\end{thm}
\begin{proof}
  *\\
\end{proof}

\subsection{State estimation}
\label{sec:state-est}

As the environment uncertainty and measurement error evolve in a normal distribution, we use a Kalman filter~\cite{stengel1994optimal, simon2006optimal} for the purpose of state estimation.

The prediction is done using
\begin{subequations}\label{eq:kalman-pred}\begin{align}
  \hat{\mathbf{q}}_{k+1}^-&=A\hat{\mathbf{q}}_{k}+B\mathbf{u}_k,\label{eq:kalman-pred1}\\
  P_{k+1}^-&=AP_kA^T+Q,\label{eq:kalman-pred2}
\end{align}\end{subequations}
where $\hat{\mathbf{q}}_k^-,\hat{\mathbf{q}}_k\in\mathbb{R}^j$ depicts the estimate of the state before and after measurement (or simply estimate), and $P_k,P_k^-\in\mathbb{R}^{j\times j}$ the error covariance matrix (i.e., the variance of the estimate).

The estimation of the state and the update of the predicted output is done using
\begin{subequations}\label{eq:kalman-est}\begin{align}
  K_k&=(CP_{k+1}^-C^T+R)^{-1}(P_{k+1}^-C^T),\\
  \hat{\mathbf{q}}_{k+1}&=\hat{\mathbf{q}}_{k+1}^-+K_k(y_k^s+y_k^c-C\hat{\mathbf{q}}_{k+1}^-),\label{eq:kalman-est2}\\
  P_{k+1}&=(I-K_kC)P_{k+1}^-,\\
  \hat{y}_k&=C\hat{\mathbf{q}}_{k+1},\label{eq:kalman-est4}
\end{align}
\end{subequations}
where $K_k\in\mathbb{R}^j$ is the gain of the Kalman filter, and $I$ the identity matrix. $y_k^s,y_k^c$ are the instantaneous energy readings: $y_k^s\in\mathbb{R}_{\geq 0}$ the robot sensor, i.e., the energy due to the trajectory, and $y_k^c$ the energy of a given software configuration described in Equation~(\ref{eq:energy-comp}). The noise covariance matrices $Q\in\mathbb{R}^{j\times j},R\in\mathbb{R}$ indicates the uncertainty and measurement error covariance respectively, and $\hat{y}_k\in\mathbb{R}_{\geq 0}$ is the estimated energy.

Equations~(\ref{eq:kalman-pred}--\ref{eq:kalman-est}) converge to the predicted energy evolution as follows. An initial guess of the values $\hat{\mathbf{q}}_0,P_0$ is derived empirically from collected data. It is worth considering that an appropriate guess of these parameters allows the algorithm to converge to the desired energy evolution in a shorter amount of time. The tuning parameters $Q,R$ are also derived from the collected data, and may differ due to i.e., different sensors used to measure the instantaneous energy consumption, or different atmospheric conditions accounting for the process noise.

At time $k=0$, the initial estimate before measurement of the state and of the error covariance matrix is updated in Equation~(\ref{eq:kalman-pred1})~and~(\ref{eq:kalman-pred2}) respectively. The value of $\hat{\mathbf{q}}_1^-$ is then used in Equation~(\ref{eq:kalman-est2}) to estimate the current state along with the data from the sensor $y_0$ (e.g., the energy sensor of the flight controller of the fixed-wing craft), where the sensor noise covariance matrix $R$ accounts for the amount of uncertainty in the measurement. The estimated output $\hat{y}_0$ is then obtained from Equation~(\ref{eq:kalman-est4}). The algorithm is iterative. At time $k=1$ the values $\hat{\mathbf{q}}_1,P_1$ computed at previous step are used to estimate the values $\hat{\mathbf{q}}_2,P_2,$ and $y_1$.

\subsection{Optimal control action}

>>>>

The overall plan is the union of all the missions. If for simplicity the system is sampled discrete-time
\begin{equation}
  \mathcal{M}:=\bigcup_{i\in[t_f]}{\mathcal{M}_i},
\end{equation}
the algorithm inputs $\mathcal{M}$ and outputs the position, the instantaneous energy consumption, and the controls sequence--an action performed evolving the mission state.

@@

The control $\mathbf{u}$ along with the input matrix
\begin{equation}\label{eq:state-control}\begin{split}  
  \mathbf{u}_k&=\begin{bmatrix}g(\mathbf{s}_k)-g(\mathbf{s}_{k-1}) & \mathbf{c}_k-\mathbf{c}_{k-1}\end{bmatrix}^T,\\
  B&=\left[\begin{array}{cccc}
    1& \omega_{k,1}& \cdots& \omega_{k,\rho}  \\
     &            0&       &  \\
     &             & \ddots&  \\
     &             &       & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{s}_k$ is defined in Subsection~\ref{sec:computations-model}, $\mathbf{u}_k\in\mathbb{R}^l$ is the control given $l:=1+\rho$, $\mathbf{u}_{1}=\begin{bmatrix}0 & \cdots & 0\end{bmatrix}^T$ and $B\in\mathbb{R}^{j\times l}$. Moreover, the first item is one, while the others on the first row are gain factors $\omega_{k}\in\mathbb{R}$, quantifying the contribution of a given adjustment to the instantaneous energy.

@@

In the case of the system behaving ideally (i.e., with no uncertainty), we expect a state (energy evolution) evolving accordingly to its output (instantaneous energy consumption). Such observation is summarized in the following Lemma.

\begin{lem}[State, output proportionality]\label{lem:state-vs-energy}
  Suppose the system of Equation~(\ref{eq:state-perf}) evolves with no uncertainty ($\mathbf{w}=\mathbf{0},v=0$) and Assumption~\ref{assm:periodic} holds. Given two time instants $k_1,k_2\in[t_f]$
  \begin{equation}
    \|\mathbf{q}_{k_1}\|\geq\|\mathbf{q}_{k_2}\|\iff y_{k_1}\geq y_{k_2}.
  \end{equation}
\end{lem}
\begin{proof}
  \begin{center}\vspace{.2cm}``\emph{The easy proof is trivial and is left as an exercise to the reader :P}''\vspace{.6cm}\end{center}
\end{proof}

@@

The plan $\mathcal{M}$ contains a set of ordered lists with $\sigma$ tasks each (recall Definition~\ref{def:mission}). These tasks are simulated by $\sigma$ ROS nodes $\Psi(\mathbf{s}_k):=\left(\psi_1(s_{k,1}),\dots,\psi_\sigma(s_{k,\sigma})\right)$, in fact they input the desired and output the actual computations. Let us define the computations
\begin{equation}\label{eq:qos-def}
  \mathbf{s}_k:=\left\{s_{k,1},\dots ,s_{k,\sigma}\mid \psi_i(s_{k,i})\in\mathbb{S}_{k}\,\forall i\in[\sigma]\right\},
\end{equation}
where $s_{k,i}:\mathbb{Z}_{\geq 0}\rightarrow\mathbb{Z}_{\geq 0}$ returns the $i$-th computation, $\mathbf{s}_k\in\mathbb{S}_k\subseteq\mathbb{Z}_{\geq 0}^\sigma$ the set of $\sigma$ computations at time $k$ being $\mathbb{S}_k:=\bigcup_{i\in[\sigma]}{\mathbb{S}_{k,i}}$.

@@

%%%%%%%%%%%%%%%%%%%
\section{Algorithm}
\label{sec:algo}

@@

Given an initial mission $\mathcal{M}_1$ (TEE $\varphi_1$, tasks $\Psi$, computations $\mathbf{s}_1\in\mathbb{S}_1$, and adaptations $\mathbf{c}_1\in\mathbb{C}_1$), the algorithm produce a valid mission evolution.

\begin{defn}[Valid mission evolution]\label{def:valid}
  At a generic time instant $k\in[t_f]$ a mission evolution $\mathcal{M}_{k}$ is valid
  \begin{equation}\begin{split}
    \mathbf{u}^a_{k}=\{(\mathbf{c}_{k},\mathbf{s}_{k})\mid&(\varphi(\mathbf{p}_{k-1},\mathbf{c}_{k-1}),\Psi(\mathbf{s_{k-1}}))\in\mathcal{M}_{k-1}\\
    &\Longrightarrow(\varphi(\mathbf{p}_{k},\mathbf{c}_{k}),\Psi(\mathbf{s}_{k}))\in\mathcal{M}_{k}\}.
  \end{split}\end{equation}
\end{defn}

Note that the position evolution can be computed directly from Equation~(\ref{eq:pd}). If the velocity is $v\in\mathbb{R}_{\geq 0}$, and starting point $\mathbf{p}_0$, $\mathbf{p}_{k+1}=\mathbf{p}_k\dot{\mathbf{p}}_d(\mathbf{p}_k)/v$.

Let us proof that if the mission evolution is valid, the instantaneous energy consumption is a linear combination of the state from the Equation~(\ref{eq:state-perf}).

\begin{thm}[State output linearity]\label{thm:state-vs-energy}
  Consider the mission from Definition~\ref{def:mission}, the valid mission evolution from~\ref{def:valid}, and assume Assumption~\ref{assm:periodic} holds. Likewise in Lemma~\ref{lem:state-vs-energy}, the model behaves ideally. 
  Then, the instantaneous energy consumption $y_k$ is a linear combination of the state $\mathbf{q}_k$ and $\mathcal{M}_k$ produce a valid mission evolution $\mathcal{M}_{k+1}\,\forall k\in[t_f-1]$.
\end{thm}
\begin{proof}
  *\\
\end{proof}

%\subsection{State estimation}
%\label{sec:state-est}

%As the environment uncertainty and measurement error evolve in a normal distribution, we use a Kalman filter~\cite{stengel1994optimal, simon2006optimal} for the purpose of state estimation.

%The prediction is done using
%\begin{subequations}\label{eq:kalman-pred}\begin{align}
%  \hat{\mathbf{q}}_{k+1}^-&=A\hat{\mathbf{q}}_{k}+B\mathbf{u}_k,\label{eq:kalman-pred1}\\
%  P_{k+1}^-&=AP_kA^T+Q,\label{eq:kalman-pred2}
%\end{align}\end{subequations}
%where $\hat{\mathbf{q}}_k^-,\hat{\mathbf{q}}_k\in\mathbb{R}^j$ depicts the estimate of the state before and after measurement (or simply estimate), and $P_k,P_k^-\in\mathbb{R}^{j\times j}$ the error covariance matrix (i.e., the variance of the estimate).

%The estimation of the state and the update of the predicted output is done using
%\begin{subequations}\label{eq:kalman-est}\begin{align}
%  K_k&=(CP_{k+1}^-C^T+R)^{-1}(P_{k+1}^-C^T),\\
%  \hat{\mathbf{q}}_{k+1}&=\hat{\mathbf{q}}_{k+1}^-+K_k(y_k^s+y_k^c-C\hat{\mathbf{q}}_{k+1}^-),\label{eq:kalman-est2}\\
%  P_{k+1}&=(I-K_kC)P_{k+1}^-,\\
%  \hat{y}_k&=C\hat{\mathbf{q}}_{k+1},\label{eq:kalman-est4}
%\end{align}
%\end{subequations}
%where $K_k\in\mathbb{R}^j$ is the gain of the Kalman filter, and $I$ the identity matrix. $y_k^s,y_k^c$ are the instantaneous energy readings: $y_k^s\in\mathbb{R}_{\geq 0}$ the robot sensor, i.e., the energy due to the trajectory, and $y_k^c$ the energy of a given software configuration described in Equation~(\ref{eq:energy-comp}). The noise covariance matrices $Q\in\mathbb{R}^{j\times j},R\in\mathbb{R}$ indicates the uncertainty and measurement error covariance respectively, and $\hat{y}_k\in\mathbb{R}_{\geq 0}$ is the estimated energy.

%Equations~(\ref{eq:kalman-pred}--\ref{eq:kalman-est}) converge to the predicted energy evolution as follows. An initial guess of the values $\hat{\mathbf{q}}_0,P_0$ is derived empirically from collected data. It is worth considering that an appropriate guess of these parameters allows the algorithm to converge to the desired energy evolution in a shorter amount of time. The tuning parameters $Q,R$ are also derived from the collected data, and may differ due to i.e., different sensors used to measure the instantaneous energy consumption, or different atmospheric conditions accounting for the process noise.

%At time $k=0$, the initial estimate before measurement of the state and of the error covariance matrix is updated in Equation~(\ref{eq:kalman-pred1})~and~(\ref{eq:kalman-pred2}) respectively. The value of $\hat{\mathbf{q}}_1^-$ is then used in Equation~(\ref{eq:kalman-est2}) to estimate the current state along with the data from the sensor $y_0$ (e.g., the energy sensor of the flight controller of the fixed-wing craft), where the sensor noise covariance matrix $R$ accounts for the amount of uncertainty in the measurement. The estimated output $\hat{y}_0$ is then obtained from Equation~(\ref{eq:kalman-est4}). The algorithm is iterative. At time $k=1$ the values $\hat{\mathbf{q}}_1,P_1$ computed at previous step are used to estimate the values $\hat{\mathbf{q}}_2,P_2,$ and $y_1$.

%\subsection{Optimal control action}

%*
